---
title: "EDH people: extraction of the dataset and basic exploration"
author: 
- Petra Hermankova^[Aarhus University, Denmark, https://orcid.org/0000-0002-6349-0540]
date: "`r format(Sys.Date())`"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 3
    df_print: paged
---

_Purpose of this script is to create a smaller subsection of the full [Epigraphic Database Heidelberg (**EDH**) dataset](https://zenodo.org/record/4888168/) that can be further explored for the purpose of the SNA within the **Past Social Network Project**, Aarhus University. A basic exploration of useful attributes connected to people on inscriptions, such as gender, age, social status, is included._


**WARNING!** The following code was designed for the 2021 version. As some attributes in the 2022 version changed, you may have to alter the code below where necessary.


# Initial setup

```{r setup, echo=TRUE, message=FALSE, warning = FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)
library(jsonlite)
library(stringi)
```

```{r dowloading data}
# getwd() # check you working directory
dir.create("../data")
dir.create("../data/large_data")

# download as a local copy from Zenodo
# version 2021 (81476)
# download.file("https://zenodo.org/record/4888168/files/EDH_text_cleaned_2021-01-21.json?download=1", "../data/large_data/EDH_text_cleaned_2021-01-21.json")

########### WARNING! In case you get an error message, you may have to manually download the dataset from Zenodo and save it to the 'large_data' folder. Don't forget to check if the dataset or the entire folder is in your .gitignore file, otherwise you are risking accidental commit to GitHub and than having to deal with removing the large file from your commit. 

# version 2022 (81883)
# download.file("https://zenodo.org/record/7303886/files/EDH_text_cleaned_2022-11-03.json?download=1", "../data/large_data/EDH_text_cleaned_2022-11-03.json")


########### WARNING! The following code was designed for the 2021 version. As some attributes in the 2022 version changed, you may have to alter the code below where necessary.
```

```{r loading data}
# 2021
list_json <- jsonlite::fromJSON("../data/large_data/EDH_text_cleaned_2021-01-21.json")

# 2022
#list_json <- jsonlite::fromJSON("../data/large_data/EDH_text_cleaned_2022-11-03.json")

EDH <- as_tibble(list_json)
```


# Subset the dataset to the minimal useful version

```{r dataset filtering}
EDHs<- EDH %>% 
  select(id,
         coordinates, findspot_ancient_clean, province_label_clean, modern_region_clean,
         not_before, not_after,
        type_of_monument_clean, material_clean,
    type_of_inscription_clean,
    transcription, clean_text_interpretive_word, 
    people
    ) %>% 
  separate(col = coordinates, into = c("latitude", "longitude"), sep = ",")

EDHs$longitude <- as.numeric(str_replace(EDHs$longitude, pattern = "\\)", replacement=""))
EDHs$latitude <- as.numeric(str_replace(EDHs$latitude, pattern = "c\\(", replacement=""))
```


```{r extracting people}
EDHs$people <- map(EDHs$people, as.data.frame)
EDH_people<- EDHs %>% unnest_longer(col = people, keep_empty = TRUE)
EDH_people<- EDH_people %>% 
  unnest_wider(people)

nrow(EDH_people)

```

```{r}
#extracting only inscriptions with people in them

EDH_people <- EDH_people %>% 
  filter(!is.na(name))

nrow(EDH_people)
```


```{r display few records}
head(EDH_people, 10)
```


# Data exploration & streamlining

## People on inscriptions

### How many people there are per inscription (average, min, max)
```{r how many people}
summary(as.numeric(EDH_people$person_id))
```

## Gender

### What is the gender ratio of people on inscriptions? (male, female, NA)
```{r gender}
EDH_people %>% 
  count(gender, sort = TRUE) %>% 
  mutate(ratio = round(n/(sum(n)/100),2))
```

## Status

### What are the names of unique values in the 'status' attribute?
```{r status}
EDH_people$status %>% 
  unique()
```


### What is the ratio of different statuses, e.g. slave vs freedman

```{r processing status}
str_split_fixed(EDH_people$status, ";", n=3) %>% 
  as.data.frame() -> status

status %>% 
  cbind(combined = c(status$V1,status$V2,status$V3)) %>% 
  filter(combined != "") %>% 
  mutate(combined_clean = str_replace_all(string = combined, pattern = "\\?", replacement = "")) %>% 
  mutate(combined_clean = str_replace_all(string = combined_clean, pattern = "^ ", replacement = "")) %>% 
  count(combined_clean, sort=TRUE) -> status_counts

status_counts 
```

```{r status figures}
status_counts %>% 
  mutate(combined_clean = reorder(combined_clean, n)) %>% 
  ggplot(aes(y=combined_clean, x=n, fill=combined_clean)) +
  geom_col(width=0.8, stat="identity") +
  coord_cartesian(xlim=c(0,10000)) +
  labs(x = "Number of instances", y = "Status category", title = "Overview of status references in the EDH dataset", subtitle = ggtitle(paste("n =", nrow(EDHs), "inscriptions"))) +
  geom_label(aes(label= n)) +
  theme_linedraw(base_size = 13) +
  theme(legend.position = "none")


dir.create("../figures")
ggsave("../figures/Status_overview.jpg", width = 12, height = 8) 
ggsave("../figures/Status_overview.png", width = 12, height = 8) 
```

## Age

Before we can display the age of people, we need to streamline the age as it is recorded as character and distributed over several attributes, such as age in years, months, days. Then convert them to numeric value and save as total age in years.

```{r processing age}
EDH_people<- EDH_people %>% 
  mutate(age_years = as.numeric(str_extract(EDH_people$'age: years', pattern = "[:digit:]+"))) %>% 
  mutate(age_months = as.numeric(str_extract(EDH_people$'age: months', pattern = "[:digit:]+"))) %>%
  mutate(age_days = as.numeric(str_extract(EDH_people$'age: days', pattern = "[:digit:]+"))) %>%
  mutate(age_hours = as.numeric(str_extract(EDH_people$'age: hours', pattern = "[:digit:]+"))) %>% 
  mutate(months_to_years = age_months / 12) %>% 
  mutate(days_to_years = age_days / 365) %>% 
  mutate(hours_to_years = age_hours / (24*365)) %>% 
  replace_na(list(months_to_years = 0, days_to_years = 0, hours_to_years = 0)) %>% 
  mutate(total_age = age_years + months_to_years + days_to_years + hours_to_years) %>% 
  dplyr::select(-ends_with("to_years")) %>% 
  dplyr::select(-starts_with("age: "))

```

### Summary of age in years

```{r total age summary}
summary(EDH_people$total_age)
```

### How many percent of people state their age on inscriptions
```{r age on inscriptions}
length(na.omit(EDH_people$total_age))/(nrow(EDH_people)/100)
```


## Origin

Before we can display the origin of people, we need to streamline the text and clean it from uncertainty symbols and other variations. That way, we can get as close to the original place name as possible.

Desription from the EDH website:   

```
Geographical information concerning the origin of a person named in the inscription

spelling guide:
usually as it is in the text of the inscription
(e.g. "Alba Pompeia", "natione Surus domo Hemesa", "ex Italia / τῶν ἐξ Ἰταλίας", "ex provincia Asia civitate Focia")
in the nominative only when adjectives are used (e.g. "Ulisitanus", "Emeritensis", "Ῥωμαῖος")
correct spelling without brackets
Designation of resolutions and supplements / erasures:
* = resolved abbreviation
+ = supplied
++ = erased, but still readable
+++ = erased and no longer readable
```


```{r cleaning origo}
EDH_people <- EDH_people %>%
  mutate(origo_clean = str_replace_all(origo, pattern="[\\*|\\+|\\?|\\!]", replacement="")) %>%
  mutate(origo_clean = str_replace_all(origo_clean, pattern="(\\w+) \\(= (\\w+)\\)", replacement="\\2")) %>%
  mutate(origo_clean = str_replace_all(origo_clean, pattern="gente|genitus|gentis|natus|civis|Civis|civus|natione|nato|regione|domo|domu|cives|cive|civi|ex|verna|tate", replacement="")) %>%
  mutate(origo_clean = str_replace_all(origo_clean, pattern="^ ", replacement="")) %>%
  mutate(origo_clean = str_replace_all(origo_clean, pattern="  ", replacement=" ")) 
```

### Summary of origo
```{r origo overview}
table(EDH_people$origo_clean) %>% 
  sort(decreasing = TRUE) %>% 
  as.data.frame()
```

## Names and family ties

### Name

Each individual person is identified on the basis of the 'name' attribute.
Here is the attribute description as found at the EDH website:

```
Name (including filiation), Tribus and Origo of a person named in the inscription

spelling:
as found and preserved in the inscriptio
include supplements, but without resolution of abbreviations
Emperors: the nomenclature "Imperator Caesar ... Augustus" is only indicated in the database for the Julian / Claudian emperors; "without Epithets (Pius, Felix, Invictus) or victory names (Parthicus maximus etc.)
```

#### Cleaning of the name
```{r cleaning name}

restoration_interpretive <- c("[\\[*\\]]", "")
substitution_edh_interpretive <- c("([α-ωΑ-Ωa-zA-Z])=([α-ωΑ-Ωa-zA-Z])", "\\1")
#substitution_interpretive <- c("[\\<*\\>]", "")
arabic_numerals <- c("[0-9]+", "")
erasure_new_text <- c("[〚〛]", "")
#erasure_new_text_triple <- c("\\[\\[\\[|\\]\\]\\]", "")
#erasure_new_text_double <- c("\\[\\[|\\]\\]", "")

cleaning_interpretive_word_edh <- function(epigraphic_dataset){      
  clean_text <- gsub(pattern=substitution_edh_interpretive[1], replacement=substitution_edh_interpretive[2], x=epigraphic_dataset, perl=TRUE)
  #clean_text <- gsub(pattern=substitution_interpretive[1], replacement=substitution_interpretive[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=erasure_new_text[1], replacement=erasure_new_text[2], x=clean_text, perl=TRUE)
  #clean_text <- gsub(pattern=erasure_new_text_triple[1], replacement=erasure_new_text_triple[2], x=clean_text, perl=TRUE)
  #clean_text <- gsub(pattern=erasure_new_text_double[1], replacement=erasure_new_text_double[2], x=clean_text, perl=TRUE)
  clean_text <- gsub(pattern=arabic_numerals[1], replacement=arabic_numerals[2], x=clean_text, perl=TRUE)
    return(clean_text)
}

EDH_people<- EDH_people %>% 
  mutate(name_clean = cleaning_interpretive_word_edh(EDH_people$name))
```

```{r test cleaning}
input_string <- "[[[P. Septimius Geta]]]"
input_string <- "[[Geta]]"
input_string <- "[[---]]"
result <- gsub("\\[\\[\\[|\\]\\]\\]", "", input_string, perl=TRUE)
result <- gsub("\\[\\[|\\]\\]", "", input_string, perl=TRUE)
cat(result)
```



```{r clean name}
EDH_people %>% 
  count(name_clean, name, sort=F) %>% 
  View()
```

Some people do have the attribute 'name' empty, but there are still relevant information. However, for the SNA they should be removed, otherwise they will create connections and relationships where there were none.

```{r}
# review inscriptions with people with an empty name in [---], [[---]], [[[---]]]

empty_name_string <- c("[---]", "[[---]]", "[[[---]]]")

# remove people with no name, use at caution
EDH_people_rn<- EDH_people %>% 
  filter(!name_clean %in% empty_name_string)
```


### Family name (nomen)

Description as found at the EDH website:

```
Nomen / Nomina of a person named in the inscription. 

Here one can search for the occurrence of a nomen independent of case, spelling and state of preservation on the inscription bearer.

Spelling:
- Always in the nominative irrespective of the case in the inscription.
- correct spelling without brackets
- In the case of the emperors up to and including Nero Caesar is treated as a part of the name.


Resolution of abbreviations and supplements / erasures are not indicate by round or square brackets ( ), [ ], [[ ]], [[[ ]]], but through:
* = resolved
+ = supplemented
++ = erased, but still readable
+++ = erased and no longer readable

```

```{r nomen exploration}
EDH_people %>% 
  count(nomen, sort=T)
```

#### Cleaning of the family name for our purposes

```{r cleaning nomen}

EDH_people<- EDH_people %>% 
  mutate(nomen_clean = str_replace_all(EDH_people$nomen, "\\+{1,3}", "")) 

EDH_people<- EDH_people %>% 
  mutate(nomen_clean = str_replace_all(EDH_people$nomen_clean, "\\*", "")) 

EDH_people<- EDH_people %>% 
  mutate(nomen_clean = str_replace_all(EDH_people$nomen_clean, "[\\+\\*]", "")) 

EDH_people<- EDH_people %>% 
  mutate(nomen_clean = str_replace_all(EDH_people$nomen_clean, "[\\!]", "")) 

EDH_people<- EDH_people %>% 
  mutate(nomen_clean = str_replace_all(EDH_people$nomen_clean, "\\)\\?", "\\)")) 

EDH_people<- EDH_people %>% 
  mutate(nomen_clean = str_replace_all(EDH_people$nomen_clean, "\\?", "")) 
```

```{r encoding conversion}
library(stringi)

# Convert the encoding of the nomen_clean column
 EDH_people$nomen_clean <- iconv(EDH_people$nomen_clean, from = "UTF-8", to = "windows-1252")
 EDH_people$nomen_clean <- iconv(EDH_people$nomen_clean, from = "windows-1252", to = "UTF-8")
 #EDH_people$name <- iconv(EDH_people$name, from = "UTF-8", to = "windows-1252")
 #EDH_people$name <- iconv(EDH_people$name, from = "windows-1252", to = "UTF-8")

```

```{r clean nomen overview}
EDH_people %>% 
  count(nomen_clean, sort=T)
```



### Indexing people

```{r, index number for people}

# creating a unique ID number for people
EDH_people$id_people <- 1:nrow(EDH_people)

# renaming variable describing the number of people on one inscription
EDH_people <-rename(EDH_people, person_number_insc = person_id)

```



Ordering the attributes so they are ordered logically
```{r ordering}
EDH_people<- EDH_people %>% 
  select(id, latitude, longitude, findspot_ancient_clean, province_label_clean, modern_region_clean, not_before, not_after, type_of_monument_clean, material_clean, type_of_inscription_clean, transcription, clean_text_interpretive_word, id_people, person_number_insc, name, praenomen, nomen, name_clean, nomen_clean, cognomen, supernomen, gender, origo, status, total_age, age_years, age_months, age_months, age_hours)
```


# Save the streamlined data as CSV
```{r save CSV}
dir.create("../data")
write.csv(EDH_people, "../data/EDH_people_2021.csv", row.names=TRUE, sep = ";")

```

# Networks potential 

## Inscriptions with 100 and more people
```{r people 100}

EDH_people %>% 
  filter(person_number_insc >100) %>% 
  count(id, clean_text_interpretive_word, sort=T) -> inscr_100_people
inscr_100_people
```

1. HD032316
https://edh.ub.uni-heidelberg.de/edh/inschrift/HD032316 inscription with 244 veterans of legio VII Claudia, separated by cohors, from Viminacium (Serbia), 195 AD, some of them have a geographic origin, e.g. Pautalia, Scupis, Remesiana, Ratiaria, Sarmizegetusa, Montana, Salonika, Nikomedia, Pergamon, Sirmii, ...

More background: 264 inscriptions from Viminacium, total 866 people

```{r} 
# filter for people from Viminacium, where not before date is bigger than 170 and smaller than 220
EDH_people %>% 
  filter(findspot_ancient_clean == "Viminacium") -> viminacium

viminacium %>% 
  filter(as.numeric(not_before) > 170 & as.numeric(not_before) < 220)
```

```{r}
View(viminacium)
```



2. HD043480
https://edh.ub.uni-heidelberg.de/edh/inschrift/HD043480 inscription with 237 veterans of unknown unit, separated by cohors, Troesmis (Romania), 134 AD. No geographic origin, but some functions.

3. HD005647
https://edh.ub.uni-heidelberg.de/edh/inschrift/HD005647, list of 163 men and women with some occupations, Classis (Italy), 171-230 AD

4. HD017350
https://edh.ub.uni-heidelberg.de/edh/inschrift/HD017350, list of 149 soldiers, names of the praetorians and legionnaires and the auxiliary soldiers, sometimes with their origin and rank, Tropaeum Traiana (Romania), 87 AD – 90 AD

5. HD000674	
https://edh.ub.uni-heidelberg.de/edh/inschrift/HD000674, list of 147 members of collegium centonarium, Flavia Solva (Austria), 205 AD

6. HD042833	
https://edh.ub.uni-heidelberg.de/edh/inschrift/HD042833, list of 142 male members of the cult of Bacchus, some of them hacve social status specified, Nicopolis ad Istrum (Bulgaria), 227 AD


7. HD065217	
https://edh.ub.uni-heidelberg.de/edh/inschrift/HD065217, list of 119 male members of Sarapis and Isis cult with their functions, Maroneia (Greece), 200 BC - 1 BC


8. HD019742	
https://edh.ub.uni-heidelberg.de/edh/inschrift/HD019742, list of 114 soldiers made citizens, also known as Decretum Cn. Pompei Strabonis de Civitate Equitibus Hispanis Danda, Roma (Italy), 89 BC

9. HD005863	
https://edh.ub.uni-heidelberg.de/edh/inschrift/HD005863, list of 103 soldiers, Classis (Italy), 151 – 230 AD


10. HD051713
https://edh.ub.uni-heidelberg.de/edh/inschrift/HD051713, list of men, some of them called pater, some of them obitus/deceased, Building/foundation inscription from Municipium Claudium Virunum (Austria), 182 - 184 AD 


## Women on inscriptions

```{r}
EDH_people %>% 
  filter(gender == "female") -> EDHfem
```

```{r}
EDHfem %>% 
  count(findspot_ancient_clean, sort = T)
```

```{r} 
# females from ancient Salona in Dalmatia

EDHfem %>%
  filter(findspot_ancient_clean %in% c("Salonae", "Salonae - 0a1")) -> SalonaFEM
```


```{r}
SalonaFEM
```

```{r} 
SalonaFEM %>% 
  count(nomen_clean, sort=T)

```

```{r}
SalonaFEM %>% 
  count(total_age, sort=T)
```


```{r}
# network from one inscription HD063871, 41 AD – 70 AD, Salonae, 14 people

# TEXT
# C(aius) Iulius Sceptus Admeti Aug(usti) lib(erti) [l]ib(ertus) IIIIIIvir Augustal(is) v(ivus) f(ecit) sibi et / Iuliae Coetonidi uxori / C(aio) Iulio C(ai ) f(ilio) Tro(mentina) Admeto f(ilio) / C(aio) I(ulio) C(ai) f(ilio) Tro(mentina) Aquilae f(ilio) / L(ucio) Iulio C(ai) f(ilio) Tro(mentina) Scepto f(ilio) / Iuliae C(ai) f(iliae) Tro(mentina) Admetidi f(iliae) // C(aio) Iulio |(mulieris) l(iberto) Tr[iu ]mphali f(ilio) / Iuliae C(ai) l(ibertae) Ro[m]anae f(iliae) / Iuliae |(mulieris) l(ibertae) S[ce]psidi f(iliae) / Iuliae |(mulieris) l(ibertae) Pr[im]ae uxori / defun[cta]e // Niso Ti(beri) Claudi Aug(usti) ser(vo) genero / defuncto / in h(oc) m(onumentum) veto aliter ossua / deponi quam Primae et nisi sunt // in fronte p(edes) XX i[na]gro p(edes) XXX



```

## Merchants on inscriptions

```{r}
EDH_people %>% 
  filter(clean_text_interpretive_word == str_subset(clean_text_interpretive_word, "empor*")) 
```

