---
title: "EDH people: exploration of the dataset"
author: 
- Petra Hermankova^[Aarhus University, Denmark, https://orcid.org/0000-0002-6349-0540]
date: "`r format(Sys.Date())`"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 2
    df_print: paged
---

## Initial setup

```{r setup, echo=TRUE, message=FALSE, warning = FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)
library(jsonlite)
```

```{r dowloading data}
getwd()
dir.create("../data")
dir.create("../data/large_data")

# download (manually) as a local copy from Zenodo
# version 2021 (81476)
# download.file("https://zenodo.org/record/4888168/files/EDH_text_cleaned_2021-01-21.json?download=1", "../data/large_data/EDH_text_cleaned_2021-01-21.json")

# version 2022 (81883)
# download.file("https://zenodo.org/record/7303886/files/EDH_text_cleaned_2022-11-03.json?download=1", "../data/large_data/EDH_text_cleaned_2022-11-03.json")
```

```{r loading data}
# 2021
list_json <- jsonlite::fromJSON("../data/large_data/EDH_text_cleaned_2021-01-21.json")

# 2022
#list_json <- jsonlite::fromJSON("../data/large_data/EDH_text_cleaned_2022-11-03.json")

EDH <- as_tibble(list_json)
```


## Subset the dataset to the minimal useful version

```{r dataset filtering}
EDHs<- EDH %>% 
  select(id,
         coordinates, findspot_ancient_clean, province_label_clean, modern_region_clean,
         not_before, not_after,
        type_of_monument_clean, material_clean,
    type_of_inscription_clean,
    transcription, clean_text_interpretive_word, 
    people
    ) %>% 
  separate(col = coordinates, into = c("latitude", "longitude"), sep = ",")

EDHs$longitude <- as.numeric(str_replace(EDHs$longitude, pattern = "\\)", replacement=""))
EDHs$latitude <- as.numeric(str_replace(EDHs$latitude, pattern = "c\\(", replacement=""))
```


```{r extracting people}
EDHs$people <- map(EDHs$people, as.data.frame)
EDH_people<- EDHs %>% unnest_longer(col = people, keep_empty = TRUE)
EDH_people<- EDH_people %>% 
  unnest_wider(people)
```

## Data exploration

**How many people there are per inscription (average, min, max)**
```{r}
summary(as.numeric(EDH_people$person_id))
```

### Gender

**What is the gender ratio of people on inscriptions? (male, female, NA)**
```{r}
EDH_people %>% 
  count(gender, sort = TRUE) %>% 
  mutate(ratio = round(n/(sum(n)/100),2))
```

### Status

**What are the names of unique values in the 'status' attribute?**
```{r}
EDH_people$status %>% 
  unique()
```


**What is the ratio of different statuses, e.g. slave vs freedman**

```{r}
str_split_fixed(EDH_people$status, ";", n=3) %>% 
  as.data.frame() -> status

status %>% 
  cbind(combined = c(status$V1,status$V2,status$V3)) %>% 
  filter(combined != "") %>% 
  mutate(combined_clean = str_replace_all(string = combined, pattern = "\\?", replacement = "")) %>% 
  mutate(combined_clean = str_replace_all(string = combined_clean, pattern = "^ ", replacement = "")) %>% 
  count(combined_clean, sort=TRUE) -> status_counts

status_counts 
```

```{r}
status_counts %>% 
  mutate(combined_clean = reorder(combined_clean, n)) %>% 
  ggplot(aes(y=combined_clean, x=n, fill=combined_clean)) +
  geom_col(width=0.8, stat="identity") +
  coord_cartesian(xlim=c(0,10000)) +
  labs(x = "Number of instances", y = "Status category", title = "Overview of status references in the EDH dataset", subtitle = ggtitle(paste("n =", nrow(EDHs), "inscriptions"))) +
  geom_label(aes(label= n)) +
  theme_linedraw(base_size = 13) +
  theme(legend.position = "none")


dir.create("../figures")
ggsave("../figures/Status_overview.jpg", width = 12, height = 8) 
ggsave("../figures/Status_overview.png", width = 12, height = 8) 
```
### Age

```{r}
EDH_people<- EDH_people %>% 
  mutate(age_years = as.numeric(str_extract(EDH_people$'age: years', pattern = "[:digit:]+"))) %>% 
  mutate(age_months = as.numeric(str_extract(EDH_people$'age: months', pattern = "[:digit:]+"))) %>%
  mutate(age_days = as.numeric(str_extract(EDH_people$'age: days', pattern = "[:digit:]+"))) %>%
  mutate(age_hours = as.numeric(str_extract(EDH_people$'age: hours', pattern = "[:digit:]+"))) %>% 
  mutate(months_to_years = age_months / 12) %>% 
  mutate(days_to_years = age_days / 365) %>% 
  mutate(hours_to_years = age_hours / (24*365)) %>% 
  replace_na(list(months_to_years = 0, days_to_years = 0, hours_to_years = 0)) %>% 
  mutate(total_age = age_years + months_to_years + days_to_years + hours_to_years) %>% 
  dplyr::select(-ends_with("to_years")) %>% 
  dplyr::select(-starts_with("age: "))

```

**Summary of age in years**

```{r}
summary(EDH_people$total_age)
```

**How many percent of people state their age on inscriptions**
```{r}
length(na.omit(EDH_people$total_age))/(nrow(EDH_people)/100)
```


### Origin

**Cleaning origo**
```{r}
EDH_people <- EDH_people %>%
  mutate(origo_clean = str_replace_all(origo, pattern="[\\*|\\+|\\?|\\!]", replacement="")) %>%
  mutate(origo_clean = str_replace_all(origo_clean, pattern="(\\w+) \\(= (\\w+)\\)", replacement="\\2")) %>%
  mutate(origo_clean = str_replace_all(origo_clean, pattern="gente|genitus|gentis|natus|civis|Civis|civus|natione|nato|regione|domo|domu|cives|cive|civi|ex|verna|tate", replacement="")) %>%
  mutate(origo_clean = str_replace_all(origo_clean, pattern="^ ", replacement="")) %>%
  mutate(origo_clean = str_replace_all(origo_clean, pattern="  ", replacement=" ")) 
```

```{r}
table(EDH_people$origo_clean) %>% 
  sort(decreasing = TRUE) %>% 
  as.data.frame()
```


# Save the data as CSV

```{r}
dir.create("../data")
write.csv(EDH_people, "../data/EDH_people_2021.csv", row.names=TRUE, sep = ";")

```

