---
title: "EDH people: family ties on inscriptions"
author: 
- Petra Hermankova^[Aarhus University, Denmark, https://orcid.org/0000-0002-6349-0540]
date: "`r format(Sys.Date())`"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 3
    df_print: paged
---

_Purpose of this script is to explore family ties on inscriptions from the [Epigraphic Database Heidelberg (**EDH**) dataset](https://zenodo.org/record/4888168/)  for the purpose of the SNA within the **Past Social Network Project**, Aarhus University._


**WARNING!** The following code was designed for the 2021 version. As some attributes in the 2022 version changed, you may have to alter the code below where necessary.


# Initial setup

```{r setup, echo=TRUE, message=FALSE, warning = FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)
library(jsonlite)
library(igraph)
library(readr)
```


```{r loading data}
# 2021
EDH_people <- read_csv("../data/EDH_people_2021.csv", progress = T, locale = readr::locale(encoding = "UTF-8"))

# 2022
# not implemented yet
```


# Exploring family ties


```{r filtering missing or uncertain nomen}

missing_nomen_regex<- "([A-Z]*)(\\(|\\[)---(\\)|\\])([a-z]*)"


EDH_people$nomen_nonempty <-gsub(pattern = missing_nomen_regex, replacement = "", x= EDH_people$nomen_clean, perl=TRUE)

EDH_people %>% 
  filter(!is.na(nomen_nonempty)) %>% 
  filter(nomen_nonempty != "[---]" & nomen_nonempty != "") -> EDH_people_nomen
```


```{r test graph}
# test with the first 20 people in the dataset
small <- head(EDH_people_nomen, 20)

small %>% 
  filter(!is.na(nomen_clean)) %>% 
  select(name, nomen_clean) -> sample


## graph from edge list

edges <- as.matrix(sample)
g <- graph_from_edgelist(edges, directed = FALSE)
plot(g)
```



```{r}
# people in the higher offices on inscriptions from Pompeii 

EDH_people_nomen %>% 
  filter(!is.na(nomen_clean)) %>% 
  filter(!is.na(name)) %>% 
  filter(!is.na(not_before)) %>%
  filter(!is.na(not_after)) %>%
  filter(!is.na(status)) %>%
  select(name, nomen_clean, findspot_ancient_clean, not_before, not_after, status, province_label_clean) -> EDH_nomen


higher_office<- str_subset(string = EDH_nomen$status, pattern = ".*higher*")

EDH_nomen_ho <- EDH_nomen %>% 
  filter(status == higher_office) 

EDH_nomen_ho_pompeii <- EDH_nomen_ho %>% 
  filter(findspot_ancient_clean == "Pompeii") 

# graph from matrix

edges_nomen_ho <- EDH_nomen_ho_pompeii %>% 
  select(name, nomen_clean) %>% 
  as.matrix()

f_nomen_ho <- graph_from_edgelist(edges_nomen_ho, directed = TRUE)
plot(f_nomen_ho)
```

```{r}
# people in the lower offices on inscriptions from Rome

lower_office<- str_subset(string = EDH_nomen$status, pattern = ".*lower*")

EDH_nomen_lo <- EDH_nomen %>% 
  filter(status == lower_office) 

EDH_nomen_lo_roma <- EDH_nomen_lo %>% 
  filter(findspot_ancient_clean == "Roma") 

# graph from edgelist

edges_nomen_lo <- EDH_nomen_lo_roma %>% 
  select(name, nomen_clean) %>% 
  as.matrix()

f_nomen_lo <- graph_from_edgelist(edges_nomen_lo, directed = FALSE)
plot(f_nomen_lo)

```

```{r}
# people in the lower offices on inscriptions from the EDH

lower_office<- str_subset(string = EDH_nomen$status, pattern = ".*lower*")

EDH_nomen_lo <- EDH_nomen %>% 
  filter(status == lower_office) 

# graph from edgelist

edges_nomen_lo <- EDH_nomen_lo %>% 
  select(name, nomen_clean) %>% 
  as.matrix(directed=FALSE)

f_nomen_lo <- graph_from_edgelist(edges_nomen_lo, directed = FALSE)
set.seed(1000)
plot(f_nomen_lo)
```

```{r}
# adjacency matrix

f_nomen_lo_mat<- igraph::as_adjacency_matrix(f_nomen_lo)

graph_lo_adjmat <- igraph::graph_from_adjacency_matrix(f_nomen_lo_mat)
set.seed(1000)
plot(graph_lo_adjmat)
```
```{r}
#ego network

ego_nets<- make_ego_graph(f_nomen_lo)

ego_nets[[2]]
set.seed(1000)
ego_size(f_nomen_lo)

# ego size of Iulius family - people holding a lower office
plot(ego_nets[[31]])

```


# Exploring types of inscriptions

```{r, funerary inscriptions}
# how many people there are on one funerary inscription

#they appear on the same inscription - there is a direct relationship between them

EDH_people %>% 
  filter(!is.na(name)) %>% 
  filter(type_of_inscription_clean == "epitaph") %>% 
  count(person_number_insc, sort=T)
```

```{r}
EDH_people %>% 
  filter(type_of_inscription_clean == "epitaph") %>% 
  count(person_number_insc, sort=T)

# transforming encoding

 EDH_people$name <- iconv(EDH_people$name, from = "UTF-8", to = "windows-1252")
 #EDH_people$name <- iconv(EDH_people$name, from = "windows-1252", to = "UTF-8")

 # filter out inscription with more than 10 family members
 
fam10 <- EDH_people %>% 
  filter(type_of_inscription_clean == "epitaph") %>% 
  filter(person_number_insc > 10)

fam10_names <- str_subset(fam10$id, "HD\\d{6}")


# turn it to network 

fam10 %>% 
  filter(id %in% fam10_names) -> network10

edges10 <- network10 %>% 
  filter(!is.na(name) & !is.na(id)) %>% 
  select(name, id) %>% 
  as.matrix()

g10 <- graph_from_edgelist(edges10, directed = FALSE)
plot(g10)


# NOTES: needs fixing of names, plus need to figure out better plotting, 

```
