---
title: "EDH people: spatial networks on inscriptions"
author: 
- Petra Hermankova^[Aarhus University, Denmark, https://orcid.org/0000-0002-6349-0540]
date: "`r format(Sys.Date())`"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 3
    df_print: paged
---

_Purpose of this script is to explore spatial relationships of people on inscriptions from the [Epigraphic Database Heidelberg (**EDH**) dataset](https://zenodo.org/record/4888168/)  for the purpose of the SNA within the **Past Social Network Project**, Aarhus University._


**WARNING!** The following code was designed for the 2021 version. As some attributes in the 2022 version changed, you may have to alter the code below where necessary.


# Initial setup

```{r setup, echo=TRUE, message=FALSE, warning = FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)
library(jsonlite)
library(igraph)
```


```{r loading data}
# 2021
EDH_people <- read_csv("../data/EDH_people_2021.csv", progress = T)

# 2022
# not implemented yet
```


# Exploring spatial relationships in the dataset

TBA

# Exploring types of inscriptions

```{r}
EDH_people %>% 
  filter(!is.na(name)) %>% 
  filter(type_of_inscription_clean == "epitaph") %>% 
  count(person_number_insc, sort=T)
```

```{r}
EDH_people %>% 
  filter(type_of_inscription_clean == "epitaph") %>% 
  count(person_number_insc, sort=T)

# filter out inscription with more than 10 family members

fam10 <- EDH_people %>% 
  filter(type_of_inscription_clean == "epitaph") %>% 
  filter(person_number_insc > 10)

fam10_names <- str_subset(fam10$id, "HD\\d{6}")


# turn it to network 

fam10 %>% 
  filter(id %in% fam10_names) -> network10

edges10 <- network10 %>% 
  filter(!is.na(name) & !is.na(id)) %>% 
  select(name, id) %>% 
  as.matrix()

g10 <- graph_from_edgelist(edges10, directed = FALSE)
plot(g10)


# NOTES: needs fixing of names, plus need to figure out better plotting
```
